\begin{abstract}
We present a unified, decision-oriented survey of \textbf{318 papers} (January 2024 -- February 2026) on embodied intelligence and world models, revealing that explicit coupling between world-model prediction and downstream control optimization is the single strongest predictor of real-world deployment robustness across the systems studied. The field has undergone a decisive transition: modular perception-planning-control pipelines are giving way to large-scale vision-language-action (VLA) policies and world-model-guided control stacks that jointly optimize representation, prediction, and decision making. We formalize the embodied control problem as a partially observable Markov decision process and derive a shared learning objective coupling latent dynamics modeling with control optimization. We propose a \textbf{three-axis taxonomy}: (1)~\emph{Functionality}---decision-coupled models that directly optimize task-facing objectives versus general-purpose models trained for broad predictive transfer; (2)~\emph{Temporal Modeling}---sequential step-by-step rollouts versus global trajectory-segment predictors; (3)~\emph{Spatial Representation}---compact latent vectors, tokenized feature sequences, and geometry-aware rendering representations. We systematize data resources and evaluation metrics across five families---task success, control stability, prediction fidelity, generalization, and compute efficiency---and compare more than 30 representative systems across eight method families under a normalized decision utility framework. A recurrent two-stage recipe (large prior followed by decision-coupled adaptation) emerges as the dominant empirical pattern. The analysis distills seven open challenges: long-horizon physical consistency, embodiment-aware representation alignment, deployment-oriented evaluation, data governance and compute efficiency, safe continual adaptation, multi-agent coordination, and interpretability/trustworthiness. We maintain a curated bibliography at \url{https://github.com/rsea2z/review-embodied}.
\end{abstract}

\noindent\textbf{Index Terms:} embodied AI, world models, vision-language-action models, robotic foundation models, long-horizon planning, autonomous driving, embodied world modeling, multi-agent coordination, interpretability.
